A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2020-12-22 cusp_j/paper/paper.tex
final modifs

\begin{abstract}
The present work introduces a new form of explicitly correlated factor in the context of the transcorrelated methods. 
The new correlation factor is obtained from the $r_{12} \approx 0$ mathematical analysis of the transcorrelated Hamiltonian, and its analytical form is obtained such that the scalar part of the effective two-electron potential reproduces the long-range interaction of the range-separated density functional theory at leading order in $1/r_{12}$. 
The resulting correlation factor is then tuned by a unique parameter $\mu$ which controls both the depth of the coulomb hole and its typical range in $r_{12}$. 
The transcorrelated Hamiltonian obtained with such a new correlation factor as a straightforward analytical expression  depending on the same parameter $\mu$, and its physical contents continuously change by varying $\mu$ : one can change from a non divergent repulsive Hamiltonian at large $\mu$ to a purely attractive one at small $\mu$. We investigate the convergence of the eigenvalues and right-eigenvectors of such new transcorrelated Hamiltonian as a function of the basis set and as a function of $\mu$ on a series of two-electron systems. 
We found that the convergence towards the complete basis set is much faster for a quite wide range values of $\mu$. We also propose a computationally cheap way to find a reasonable value of $\mu$ that automatically adapts to the system through its density.  
It should be pointed out that all two-electron integrals can be carried analytically within such approach and the three-electrons terms can be performed using a mixed analytical-numerical integration approach.

@@

\maketitle
\section{Introduction}
One of the most challenging problem in modern computational chemistry is the accurate simulation of the electronic structure of atomic and molecular systems, which implies solving the Schroedinger equation for a general $N$-body system. 
At the heart of the complexity of such a task relies the rapidly prohibitive computational cost of the wave function methods (WFT), which is induced by the exponential growth with the system size of the Hilbert space involved in the linear eigenvalue problem to be solved.

@@

Even if these selection schemes are extremely efficient to obtain a qualitative description of the wave function, a quantitative description is often out of reach as it implies the use of very large one-electron basis sets which are needed to recover the slowly converging correlation effects near the electron-electron coalescence point. 
These short-range correlation effects are mainly due to the diverging character of the coulomb potential, which induces a non analytical behaviour of the wave function at small inter-electronic distance ($r_{12}\approx 0$): the famous electronic cusp originally derived by Kato\cite{Kat-CPAM-57}. For a general and clear derivation of the cusp conditions for Coulombic systems up to second order, see Ref.\onlinecite{Tew-JCP-08}. As shown by the seminal works of Hylleraas\cite{Hyl-ZP-29} and further developed byKutzelnigg and coworkers, \cite{Kut-TCA-85,KutKlo-JCP-91, NogKut-JCP-94} the slow convergence of the correlation energy with respect to the quality of the basis set is mainly due to the impossibility to represent the cusp in a finite one-electron basis set. 
As the cusp conditions originate from the divergence of the Coulomb potential at $r_{12}=0$, an alternative approach would be to develop theories dealing with a smooth potential. Such a path was followed by three distinct branches: 
the explicitly correlated methods (F12), transcorrelated methods (TC) and the range-separated density functional theory (RS-DFT).

@@

Another approach to tackle the problems of WFT related to the electron cusp have been proposed by the F12 methods which introduce a function explicitly depending on the inter-electronic coordinate\cite{Ten-TCA-12,TenNog-WIREs-12,HatKloKohTew-CR-12, KonBisVal-CR-12, GruHirOhnTen-JCP-17, MaWer-WIREs-18} (called a Jastrow factor) to describe short-range correlation effects which are absent from the finite basis set. Thanks to strong orthogonality, 
all redundant correlation effects between the Jastrow factor and the basis set are projected out of the wave function, 
and therefore the Jastrow factor only deals with correlation effects orthonormal to the basis set. 
The resulting F12 theories, mainly based on perturbation theory and coupled cluster theory, are therefore 
extremely close to their parent WFT theory, with additional contributions involving two- and three-body terms resulting from the use of the Jastrow factor. 
It is important to notice that because of the orthogonalization between the Jastrow and the basis set, 
the complexity of the $N$-body problem within a given basis set is essentially the same between a F12 theory and the usual theory in WFT. Therefore, no strong compression of the electronic wave function is obtained within a given basis set by the F12 theory, 
but a fast convergence toward the complete basis set limit (CBS) is nevertheless obtained for correlation energies and other properties sensitive to the basis set.

An alternative point of view using a Jastrow factor is the so-called transcorrelated (TC) approach where the full effect of the Jastrow factor is incorporated into the calculation through a similarity transformation of the original Hamiltonian by the Jastrow factor. 
Seminal equations were derived by Hirschfelder\cite{Hirschfelder-JCP-63} who obtained an effective non hermitian operator based on a specific form of Jastrow factor, and latter on Boys, Handy and co-workers\cite{BoyHan-PRSLA-69,BoyHanLin-1-PRSLA-69,BoyHanLin-2-PRSLA-69} derived the equations of the general form of the TC Hamiltonian for an exponential form of the Jastrow factor. 
The obtained TC Hamiltonian contains certain new features with respect to the usual Hamiltonian: the TC Hamiltonian contains an additional effective two- and three-electron scalar potential together with is a non hermitian two-electron differential operator. Because of its non-hermitian nature, the TC Hamiltonian looses the variational principle, and Boys and Handy managed to derive equations to optimize orbitals and Jastrow parameters for a single Slater determinant. 
Nevertheless, such a constrained form of the wave function for the Slater together with the loose of variational principle makes it difficult to find a monotonic convergence of computed energies\cite{Handy-MolPhys-71}.

Later on, Ten-No\cite{TenNo-CPL-00-a} proposed to significantly change the paradigm of Boys and Handy: instead of using a single Slater determinant and optimizing the orbitals and the Jastrow factor, one uses a more elaborate many body theory with a frozen universal Jastrow factor.  This strategy was developed using as M{\o}ller-Plesset at second order (MP2) in Refs.\onlinecite{TenNo-CPL-00-a,HinTanTen-JCP-01} and a linearised coupled cluster ansatz in later work\cite{HinTanTen-CPL-02}.  
The Jastrow factor introduced by Ten-No, developed as a linear combination of gaussian functions and referred to as the frozen gaussian geminal (FROGG), only depends on the inter-electronic coordinate and is optimized such that the scalar effective two-electron  potential cancels on average the coulomb potential near $r_{12}=0$. The use of gaussian functions to represent the Jastrow factor enabled Ten-No to develop a numerical scheme to evaluate exactly the three-electron integrals present in TC Hamiltonian\cite{TenNo-CPL-00-b}, which is an alternative to the density-fitting method proposed by Ten-No in the general context of F12 methods\cite{TenMan-JCP-03}. 
In Ref.\onlinecite{HinTanTen-JCP-01}, Ten-No \textit{et. al.} introduced a biorthogonal approach, 
which allows to treat the non-hermitian TC Hamiltonian in a mathematical form more suited 
to develop many-body perturbation theory approaches. 
It is also noteworthy that the eigenfunctions of the TC Hamiltonian are invariant by orbital rotation, 
and therefore in the limit where one gives full flexibility to the wave function, 
the biorthogonal and usual orthogonal approaches give the same results. 
Nevertheless, as approximations are necessary done in the wave function ansatz for realistic systems, 
the biorthogonal approach might improve the quality of the results in practice.

@@

More recently, Alavi \textit{et. al}\cite{CohLuoGutDobTewAla-JCP-19} applied the TC equations with the use of elaborate Jastrow factors obtained from quantum Monte Carlo calculations, and explicitly taking into account for electron-electron together with electron-electron-nucleus correlation. The Jastrow factor used in such work were obtained from the optimized Jastrow factors for He to Ne of Moskowitz \textit{et. al.}\cite{SchMos-JCP-90}. 
In contrast to previous works where a constrained form was given to the wave function (such as coupled cluster for instance\cite{HinTanTen-CPL-02}),  Alavi \textit{et. al.} allow a full flexibility to the Slater part of the wave function to adapt to the elaborated Jastrow factors: they fully solve the $N$-electron problem corresponding to the TC Hamiltonian within a given basis set. 
This is done through the use of the FCI-QMC\cite{BooThoAla-JCP-09,BooAla-JCP-10,BooCleThoAla-JCP-11,GhaLozAla-JCP-19,VitAlaKat-JCTC-20} which is a projective technique allowing one to obtain the ground state of an operator, hermitian or not, through a stochastic sampling of the corresponding Hilbert space. Their approach share common points with that of Ten-No in the sense that they both use a fixed Jastrow factor and use a flexible form of wave functions, 
even though the FCIQMC is certainly more flexible than the coupled cluster ansatz of Ten-No used in Ref. \onlinecite{HinTanTen-JCP-01}.

@@

\subsection{A new form of Jastrow factor $u(\mu,r_{12})$ mimicking a long-range interaction}
\label{sec:new_j}
\subsubsection{A simple physical picture: the transcorrelated Hamiltonian for the helium atom}
\label{sec:he_htilde}
Let us write the Hamiltonian of the helium atom using the $r_i = \big|{\bf r}_i\big|$ and $ r_{12} = \big|{\bf r}_1 - {\bf r}_2 \big|$ coordinates\cite{Hylleraas-AdvQChem-64}  
\begin{equation}
 H  = h_c + \frac{1}{r_{12}},
\end{equation}
where 
\begin{equation}
 \begin{aligned}
 \label{def:h_c}
 h_c = &-\frac{1}{2} \sum_{i=1}^2 \bigg(\deriv{}{r_i}{2} + \frac{2}{r_i} \deriv{}{r_i}{} + \frac{2 Z}{r_i}\bigg) \\
     &-\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg) \\
     &-\bigg( \frac{\bd{r_1}}{r_1} \cdot \frac{\bd{r_{12}} }{r_{12}}  \deriv{}{r_1}{} + 
                \frac{\bd{r_2}}{r_2} \cdot \frac{\bd{r_{21}} }{r_{21}}  \deriv{}{r_2}{} \bigg).
 \end{aligned}
\end{equation}
\label{sec:he_j}
Now let us consider the transcorrelated Hamiltonian $\tilde{H}[u]$ obtained by a similarity transformation of the usual Hamiltonian by a Jastrow factor $u(r_{12})$ depending only on $r_{12}$: 
\begin{equation}
 \label{eq:ht_0}
 \begin{aligned}
 \tilde{H}[u]&= e^{-u(r_{12})} H e^{u(r_{12})}.
 \end{aligned}
\end{equation}
The only additional terms arising in $\tilde{H}[u]$ with respect to $H$ are those coming from the action of the differential operator in $r_{12}$,
\begin{equation}
 \mathcal{T}[u] =  -e^{-u(r_{12})}\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg)e^{u(r_{12})}.  
\end{equation}
By defining the following operators 
\begin{equation}
 \label{eq:def_tt}
 \tilde{t}[u] = -2 \deriv{u(r_{12})}{r_{12}}{} \deriv{}{r_{12}}{},
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{W}[u] = -\frac{2}{r_{12}} \deriv{u(r_{12})}{r_{12}}{}  , 
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{w}[u] = -\deriv{u(r_{12})}{r_{12}}{2} - \bigg( \deriv{u(r_{12})}{r_{12}}{} \bigg)^2, 
\end{equation}
one can write the Similarity transformed Hamiltonian  as
\begin{equation}
 \label{eq:ht_general}
 \tilde{H}[u] = h_c + \frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u] + \tilde{t}[u].
\end{equation}
Therefore, one can see from Eq. \eqref{eq:ht_general} that the two electrons experience a modified scalar potential given by $\frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u]$, and an additional differential operator $\tilde{t}[u]$, 
the latter making $\tilde{H}[u]$ non-hermitian.

@@

The main difference of the present approach with the FROGG introduced by Ten-No\cite{TenNo-CPL-00-a} is that instead of optimizing the Jastrow factor in terms of Gaussian functions such that $\frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u] \approx 0$ when $r_{12}=0$, 
we impose the analytical structure of $u(r_{12};\mu)$ such that i) the sum of all terms in $1/r_{12}$ (\textit{i.e.} $\tilde{W}[u] + \frac{1}{r_{12}}$) exactly provides the non divergent long-range interaction $\text{erf}(\mu r_{12})/r_{12}$ for all $r_{12}$, 
ii) the analytical form of $u(r_{12};\mu)$ is obtained from analytical conditions and not by a fit as in the case of the FROGG,  
iii) the new Jastrow factor $u(r_{12};\mu)$ is more flexible than the FROGG as it is tuned by a unique parameter $\mu$.

\subsubsection{A few properties of $u(r_{12};\mu)$}
The new Jastrow $u(r_{12};\mu)$ factor is tuned by a unique parameter, $\mu$, which have the unit of the inverse of a distance. 
To understand the physical content together with the mathematical properties of $u(r_{12};\mu)$, 
one must keep in mind that the complete correlating factor $\mathcal{J}(r_{12};\mu)$ is 
\begin{equation}
 \mathcal{J}(r_{12};\mu) = \text{exp}\big(u(r_{12};\mu)\big),
\end{equation}
and that the exact wave function $\psiex(\br{}_1,\br{}_2)$ is formally expressed as 
\begin{equation}
 \psiex(\br{}_1,\br{}_2) = \mathcal{J}(r_{12};\mu) \phimu(\br{}_{1},\br{}_{2}).  
\end{equation}
From Eq.\eqref{eq:def_j} we can notice that 
\begin{equation}
 \lim_{r_{12} \rightarrow \infty}u(r_{12};\mu) = 0,
\end{equation}
which means that 
\begin{equation}
 \lim_{r_{12} \rightarrow \infty}\mathcal{J}(r_{12};\mu) = 1,
\end{equation}
and therefore $u(r_{12};\mu)$ impacts only the small $r_{12}$ behaviour,  
which is expected since the equation determining $u(r_{12},\mu)$ is obtained from the small $r_{12}$ limit of the Schroedinger equation (see Eq.\eqref{def_j_00}).

To better understand the physical content of at $r_{12}\approx 0$,  
let us Taylor expand the function $u(r_{12};\mu)$ 
\begin{equation}
 \label{eq:j_dl}
 u(r_{12};\mu) = -\frac{1}{2\sqrt{\pi}\mu} + \frac{1}{2}r_{12} - \frac{\mu}{2\sqrt{\pi}} r_{12}^2 + o(r_{12}^4).
\end{equation}
A first remark is that $u(r_{12};\mu)$ exactly restores the cusp as
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0} \deriv{}{r_{12}}{}u(r_{12};\mu) = \frac{1}{2},
\end{equation}
implying that the right-eigenvector $\phimu(\br{}_{1},\br{}_{2})$ is cusp less (see Appendix \ref{sec:cusp}). 
The Taylor expansion of Eq. \eqref{eq:j_dl} inserted in $\mathcal{J}(r_{12},\mu)$ leads to 
\begin{equation}
 \label{eq:bigJ_dl}
 \mathcal{J}(r_{12};\mu) \approx e^{-\frac{1}{2\sqrt{\pi}\mu}} e^{ \frac{1}{2}r_{12}} e^{- \frac{\mu}{2\sqrt{\pi}} r_{12}^2}. 
\end{equation}
From Eq. \eqref{eq:bigJ_dl} one can analyze each term: a) as $e^{-\frac{1}{2\sqrt{\pi}\mu}}<1$, the effective wave function  $\phiex(\br{}_1,\br{}_2)$ has necessary a larger value at coalescence (\textit{i.e.} $\br{}_1=\br{}_2)$), 
b) the term in $e^{ \frac{1}{2}r_{12}} $ restores exactly the lack of cusp of $\phiex(\br{}_1,\br{}_2)$ (see Eq. \eqref{eq:cusp_phi_2}); and eventually last term reshapes the curvature of the $\phiex(\br{}_1,\br{}_2)$.

Coming now to the behaviour of $u(r_{12};\mu)$ as a function of the parameter $\mu$, we report in Figure \ref{fig_j_mu} its dependence on $r_{12}$ for a set of values of $\mu$, and we compare it with the FROGG. 
\begin{figure*}
 \label{fig_j_mu}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/small_mu_j.pdf}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/small_mu_exp_j.pdf}\\
        \caption{Shape of $u(r_{12};\mu)$ (left) and of $\mathcal{J}(r_{12};\mu) = \text{exp}\bigg(u(r_{12};\mu) \bigg) $ (right) for different values of $\mu$. 
We also compare the FROGG (see Eq. \eqref{eq_frogg}) and the value of $\mu = 0.865$ which equals the FROGG at $r_{12}=0$ (see Eq. \eqref{mu_ten_no_1}). }
\end{figure*}
From Figure\ref{fig_j_mu}, one can see that: i) the larger the $\mu$, the more short-range is $u(r_{12};\mu)$, ii) the larger the $\mu$, the closer $u(r_{12};\mu)$ is from 0 and therefore the less impact the correlating factor has on the wave function, iii) the smaller the $\mu$, the deeper and broader is the hole imposed by $\mathcal{J}(r_{12};\mu)$.

>>>>

\begin{abstract}
The present work introduces a new form of explicitly correlated factor in the context of the transcorrelated methods. 
The new correlation factor is obtained from the $r_{12} \approx 0$ mathematical analysis of the transcorrelated Hamiltonian, and its analytical form is obtained such that the scalar part of the effective two-electron potential reproduces the long-range interaction of the range-separated density functional theory at leading order in $1/r_{12}$. 
The resulting correlation factor exactly imposes the cusp and is tuned by a unique parameter $\mu$ which controls both the depth of the coulomb hole and its typical range in $r_{12}$. 
The transcorrelated Hamiltonian obtained with such a new correlation factor has a straightforward analytical expression  depending on the same parameter $\mu$, and its physical contents continuously change by varying $\mu$ : one can change from a non divergent repulsive Hamiltonian at large $\mu$ to a purely attractive one at small $\mu$. We investigate the convergence of the eigenvalues and right-eigenvectors of such new transcorrelated Hamiltonian as a function of the basis set and as a function of $\mu$ on a series of two-electron systems. 
We found that the convergence towards the complete basis set is much faster for a quite wide range values of $\mu$. We also propose a way to find a reasonable value of $\mu$ that automatically adapts to the system through its density.  
It should be pointed out that all two-electron integrals can be carried analytically within such approach and the three-electrons terms can be performed using a mixed analytical-numerical integration approach.

@@

\maketitle
\section{Introduction}
One of the most challenging problem in computational chemistry is the accurate simulation of the electronic structure of atomic and molecular systems, which implies solving the Schroedinger equation for a general $N$-body system. 
At the heart of the complexity of such a task relies the rapidly prohibitive computational cost of the wave function methods (WFT), which is induced by the exponential growth with the system size of the Hilbert space involved in the linear eigenvalue problem to be solved.

@@

Even if these schemes are extremely efficient to obtain a qualitative description of the wave function through a meaningful selection in an exponentially growing Hilbert space, a quantitative description is often out of reach as it implies the use of very large one-electron basis sets which are needed to recover the slowly converging correlation effects near the electron-electron coalescence point. 
These short-range correlation effects are mainly due to the diverging character of the coulomb potential, which induces a non analytical behaviour of the wave function at small inter-electronic distance ($r_{12}\approx 0$): the famous electronic cusp originally derived by Kato\cite{Kat-CPAM-57}. For a general and clear derivation of the cusp conditions for Coulombic systems up to second order, see Ref.\onlinecite{Tew-JCP-08}. As shown by the seminal works of Hylleraas\cite{Hyl-ZP-29} and further developed byKutzelnigg and coworkers, \cite{Kut-TCA-85,KutKlo-JCP-91, NogKut-JCP-94} the slow convergence of the correlation energy with respect to the quality of the basis set is mainly due to the impossibility to represent the cusp in a finite one-electron basis set. 
As the cusp conditions originate from the divergence of the Coulomb potential at $r_{12}=0$, an alternative approach would be to develop theories dealing with a smooth potential. Such a path was followed by three distinct branches: 
the explicitly correlated methods (F12), transcorrelated methods (TC) and the range-separated density functional theory (RS-DFT).

@@

Another approach to tackle the problems of WFT related to the electron cusp have been proposed by the F12 methods which introduce a function explicitly depending on the inter-electronic coordinate\cite{Ten-TCA-12,TenNog-WIREs-12,HatKloKohTew-CR-12, KonBisVal-CR-12, GruHirOhnTen-JCP-17, MaWer-WIREs-18} (called a Jastrow factor) to describe short-range correlation effects which are absent from the finite basis set. Thanks to strong orthogonality, 
all redundant correlation effects between the Jastrow factor and the basis set are projected out of the wave function, 
and therefore the Jastrow factor only deals with correlation effects orthonormal to the basis set. 
The resulting F12 theories, mainly based on perturbation theory and coupled cluster theory, are therefore 
extremely close to their parent WFT theory, with additional contributions involving two- and three-body terms resulting from the use of the Jastrow factor. 
It is important to notice that because of the orthogonalization between the Jastrow and the basis set, 
the complexity of the $N$-body problem within a given basis set is essentially the same between a F12 theory and the parent theory in WFT. Therefore, no strong compression of the electronic wave function is obtained within a given basis set by the F12 theory, 
but a fast convergence toward the complete basis set limit (CBS) is nevertheless obtained for correlation energies and other properties sensitive to the basis set.

An alternative point of view using a Jastrow factor is the so-called transcorrelated (TC) approach where the full effect of the Jastrow factor is incorporated into the calculation through a similarity transformation of the original Hamiltonian by the Jastrow factor. 
Seminal equations were derived by Hirschfelder\cite{Hirschfelder-JCP-63} who obtained an effective non hermitian operator based on a specific form of Jastrow factor, and latter on Boys, Handy and co-workers\cite{BoyHan-PRSLA-69,BoyHanLin-1-PRSLA-69,BoyHanLin-2-PRSLA-69} derived the equations of the general form of the TC Hamiltonian for an exponential form of the Jastrow factor. 
The obtained TC Hamiltonian contains certain new features with respect to the usual Hamiltonian: the TC Hamiltonian contains an additional effective two- and three-electron scalar potential together with is a non hermitian two-electron differential operator. Because of its non-hermitian nature, the TC Hamiltonian looses the variational principle, and Boys and Handy managed to derive equations to optimize both orbitals and Jastrow parameters for a single Slater determinant. 
Nevertheless, such a constrained form of the wave function for the Slater together with the loose of variational principle makes it difficult to find a monotonic convergence of computed energies\cite{Handy-MolPhys-71}.

Later on, Ten-No\cite{TenNo-CPL-00-a} proposed to significantly change the paradigm of Boys and Handy: instead of using a single Slater determinant and optimizing the orbitals and the Jastrow factor, one uses a more elaborate many body theory with a frozen universal Jastrow factor.  This strategy was developed using as M{\o}ller-Plesset at second order (MP2) in Refs.\onlinecite{TenNo-CPL-00-a,HinTanTen-JCP-01} and a linearised coupled cluster ansatz in later work\cite{HinTanTen-CPL-02}.  
The Jastrow factor introduced by Ten-No, developed as a linear combination of gaussian functions and referred to as the frozen gaussian geminal (FROGG), only depends on the inter-electronic coordinate and is optimized such that the scalar effective two-electron  potential cancels on average the coulomb potential near $r_{12}=0$. The use of gaussian functions to represent the Jastrow factor enabled Ten-No to develop a numerical scheme to evaluate exactly the three-electron integrals present in TC Hamiltonian\cite{TenNo-CPL-00-b}, which is an alternative to the density-fitting method proposed by the same author in the general context of F12 methods\cite{TenMan-JCP-03}. 
In Ref.\onlinecite{HinTanTen-JCP-01}, Ten-No \textit{et. al.} introduced a biorthogonal approach, 
which allows to treat the non-hermitian TC Hamiltonian in a mathematical form more suited 
to develop approaches based on many-body perturbation theory. 
It is also noteworthy that the eigenfunctions of the TC Hamiltonian are invariant by orbital rotation, 
and therefore in the limit where one gives full flexibility to the wave function, 
the biorthogonal and usual orthogonal approaches give the same results. 
Nevertheless, as approximations are necessary done in the wave function ansatz for realistic systems, 
the biorthogonal approach might improve the quality of the results in practice.

@@

More recently, Alavi \textit{et. al}\cite{CohLuoGutDobTewAla-JCP-19} applied the TC equations with the use of elaborate Jastrow factors which explicitly taking into account for electron-electron together with electron-electron-nucleus correlation. The Jastrow factor used in such work were obtained from the optimized Jastrow factors for He to Ne of Moskowitz \textit{et. al.}\cite{SchMos-JCP-90} in the context of quantum Monte Carlo. 
In contrast to previous works where a constrained form was given to the wave function (such as coupled cluster for instance\cite{HinTanTen-CPL-02}),  Alavi \textit{et. al.} allow a full flexibility to the Slater part of the wave function to adapt to the elaborated Jastrow factors: they fully solve the $N$-electron problem corresponding to the TC Hamiltonian within a given basis set. 
This is done through the use of the FCI-QMC\cite{BooThoAla-JCP-09,BooAla-JCP-10,BooCleThoAla-JCP-11,GhaLozAla-JCP-19,VitAlaKat-JCTC-20} which is a projective technique allowing one to obtain the ground state of an operator, hermitian or not, through a stochastic sampling of the corresponding Hilbert space. Their approach share common points with that of Ten-No in the sense that they both use a fixed Jastrow factor and use a flexible form of wave functions, 
even though the FCIQMC is certainly more flexible than the coupled cluster ansatz of Ten-No used in Ref. \onlinecite{HinTanTen-JCP-01}.

@@

\subsection{A new form of Jastrow factor $u(\mu,r_{12})$ mimicking a long-range interaction}
\label{sec:new_j}
\subsubsection{A simple physical picture: the transcorrelated Hamiltonian for the helium atom}
\label{sec:he_htilde}
Let us write the Hamiltonian of the helium atom using the $r_i = \big|{\bf r}_i\big|$ and $ r_{12} = \big|{\bf r}_1 - {\bf r}_2 \big|$ coordinates\cite{Hylleraas-AdvQChem-64}  
\begin{equation}
 H  = h_c + \frac{1}{r_{12}},
\end{equation}
where 
\begin{equation}
 \begin{aligned}
 \label{def:h_c}
 h_c = &-\frac{1}{2} \sum_{i=1}^2 \bigg(\deriv{}{r_i}{2} + \frac{2}{r_i} \deriv{}{r_i}{} + \frac{2 Z}{r_i}\bigg) \\
     &-\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg) \\
     &-\bigg( \frac{\bd{r_1}}{r_1} \cdot \frac{\bd{r_{12}} }{r_{12}}  \deriv{}{r_1}{} + 
                \frac{\bd{r_2}}{r_2} \cdot \frac{\bd{r_{21}} }{r_{21}}  \deriv{}{r_2}{} \bigg).
 \end{aligned}
\end{equation}
\label{sec:he_j}
Now let us consider the transcorrelated Hamiltonian $\tilde{H}[u]$ obtained by a similarity transformation of the usual Hamiltonian by a Jastrow factor $u(r_{12})$ depending only on $r_{12}$: 
\begin{equation}
 \label{eq:ht_0}
 \begin{aligned}
 \tilde{H}[u]&= e^{-u(r_{12})} H e^{u(r_{12})}.
 \end{aligned}
\end{equation}
With respect to $H$, the only additional terms arising in $\tilde{H}[u]$ are those coming from the action of the differential operator in $r_{12}$,
\begin{equation}
 \mathcal{T}[u] =  -e^{-u(r_{12})}\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg)e^{u(r_{12})}.  
\end{equation}
By defining the following operators 
\begin{equation}
 \label{eq:def_tt}
 \tilde{t}[u] = -2 \deriv{u(r_{12})}{r_{12}}{} \deriv{}{r_{12}}{},
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{W}[u] = -\frac{2}{r_{12}} \deriv{u(r_{12})}{r_{12}}{}  , 
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{w}[u] = -\deriv{u(r_{12})}{r_{12}}{2} - \bigg( \deriv{u(r_{12})}{r_{12}}{} \bigg)^2, 
\end{equation}
one can write the Similarity transformed Hamiltonian  as
\begin{equation}
 \label{eq:ht_general}
 \tilde{H}[u] = h_c + \frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u] + \tilde{t}[u].
\end{equation}
Therefore, one can see from Eq. \eqref{eq:ht_general} that the two electrons experience a modified scalar potential given by $\frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u]$, and an additional differential operator $\tilde{t}[u]$, 
the latter making $\tilde{H}[u]$ non-hermitian.

@@

The main difference of the present approach with the FROGG introduced by Ten-No\cite{TenNo-CPL-00-a} is that instead of optimizing the Jastrow factor as a linear combination of Gaussian functions such that $\frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u] \approx 0$ when $r_{12}=0$, 
we impose the analytical structure of $u(r_{12};\mu)$ such that i) the sum of all terms in $1/r_{12}$ (\textit{i.e.} $\tilde{W}[u] + \frac{1}{r_{12}}$) exactly provides the non divergent long-range interaction $\text{erf}(\mu r_{12})/r_{12}$ for all $r_{12}$, 
ii) the analytical form of $u(r_{12};\mu)$ is obtained from analytical conditions and not by a fit as in the case of the FROGG,  
iii) the new Jastrow factor $u(r_{12};\mu)$ is more flexible than the FROGG as it is tuned by a unique parameter $\mu$.

\subsubsection{A few properties of $u(r_{12};\mu)$}
The new Jastrow $u(r_{12};\mu)$ factor is tuned by a unique parameter, $\mu$, which have the unit of the inverse of a distance. 
From Eq.\eqref{eq:def_j} we can notice that 
\begin{equation}
 \lim_{r_{12} \rightarrow \infty}u(r_{12};\mu) = 0,
\end{equation}
which means that 
\begin{equation}
 \lim_{r_{12} \rightarrow \infty}e^{u(r_{12};\mu)} = 1,
\end{equation}
and therefore $u(r_{12};\mu)$ impacts only the small $r_{12}$ behaviour,  
which is expected since the equation determining $u(r_{12},\mu)$ is obtained from the small $r_{12}$ limit of the Schroedinger equation (see Eq.\eqref{def_j_00}).

To better understand the physical content of at $r_{12}\approx 0$,  
let us Taylor expand the function $u(r_{12};\mu)$ 
\begin{equation}
 \label{eq:j_dl}
 u(r_{12};\mu) = -\frac{1}{2\sqrt{\pi}\mu} + \frac{1}{2}r_{12} - \frac{\mu}{2\sqrt{\pi}} r_{12}^2 + o(r_{12}^4).
\end{equation}
A first remark is that $u(r_{12};\mu)$ exactly restores the cusp as
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0} \deriv{}{r_{12}}{}u(r_{12};\mu) = \frac{1}{2},
\end{equation}
implying that the right-eigenvectors of the TC Hamiltonian obtained with $u(r_{12};\mu)$ are cusp less (see Appendix \ref{sec:cusp}). 
The Taylor expansion of Eq. \eqref{eq:j_dl} inserted in the exponential correlating Jastrow factor reads 
\begin{equation}
 \label{eq:bigJ_dl}
 e^{u(r_{12},\mu)}\approx e^{-\frac{1}{2\sqrt{\pi}\mu}} e^{ \frac{1}{2}r_{12}} e^{- \frac{\mu}{2\sqrt{\pi}} r_{12}^2}. 
\end{equation}
From Eq. \eqref{eq:bigJ_dl} one can analyze each term: a) as $e^{-\frac{1}{2\sqrt{\pi}\mu}}<1$, Jastrow factor digs the Coulomb hole, 
b) the term in $e^{ \frac{1}{2}r_{12}} $ exactly imposes the cusp; c) and the last term $e^{- \frac{\mu}{2\sqrt{\pi}} r_{12}^2}$ reshapes the curvature of the Coulomb hole.

Coming now to the behaviour of $u(r_{12};\mu)$ as a function of the parameter $\mu$, we report in Figure \ref{fig_j_mu} its dependence on $r_{12}$ for a set of values of $\mu$, and we compare it with the FROGG. 
\begin{figure*}
 \label{fig_j_mu}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/small_mu_j.pdf}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/small_mu_exp_j.pdf}\\
        \caption{Shape of $u(r_{12};\mu)$ (left) and of $\text{exp}\big(u(r_{12};\mu) \big) $ (right) for different values of $\mu$. 
We also compare the FROGG (see Eq. \eqref{eq_frogg}) and the value of $\mu = 0.865$ which equals the FROGG at $r_{12}=0$ (see Eq. \eqref{mu_ten_no_1}). }
\end{figure*}
From Figure\ref{fig_j_mu}, one can see that: i) the larger the $\mu$, the more short-range is $u(r_{12};\mu)$, ii) the larger the $\mu$, the closer $u(r_{12};\mu)$ is from 0 and therefore the less impact the correlating factor has on the wave function, iii) the smaller the $\mu$, the deeper and broader is the hole imposed by $e^{u(r_{12},\mu)}$.

As the function $u(r_{12};\mu)$ contains a unique parameter, one can try to find the value of $\mfrogg$ such that 
$u(r_{12};\mfrogg)$ qualitatively reproduces the FROGG. 
To do so, we propose a one-point fit such that 
\begin{equation}
 \label{mu_ten_no_1}
u(r_{12}=0;\mfrogg) = u^{\text{TN}}(r_{12}=0), 
\end{equation}
where $ u^{\text{TN}}(x)$ is the FROGG introduced by Ten-No
\begin{equation}
 \label{eq_frogg}
 u^{\text{TN}}(x) = -\sum_{i=1}^6 c_i e^{-\alpha_i x^2 },
\end{equation}
with the coefficients $c_i$ and exponents $\alpha_i$ are given in Table 1 of Ref. \onlinecite{TenNo-CPL-00-a}. 
Eq. \eqref{mu_ten_no_1} gives 
\begin{equation}
 \label{mu_ten_no_2}
 \mfrogg = -\frac{1}{2\sqrt{\pi}u^{\text{TN}}(0)},
\end{equation}
which gives $\mfrogg\approx0.87$. 
We represent on Fig \ref{fig_j_mu} the comparison between $u(r_{12};\mfrogg)$ and the FROGG, and it is clearly apparent  that $u(r_{12};\mfrogg)$ reproduces almost perfectly the FROGG, although it decays slightly faster at large $r_{12}$. Therefore, one can see the FROGG as a special case of $u(r_{12};\mu)$  with $\mu = \mfrogg$.
